---
title: "Empirical_Bayes"
author: "Mi Zhang"
date: "5/09/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse,gridExtra,kableExtra, ggplot2, cowplot,deconvolveR)
```

# European automobile insurance
```{r warning=FALSE}
Claims <- seq(0, 7)
counts <- c(7840, 1317, 239, 42, 14, 4, 4, 1)
ins <- data.frame(Claims, counts)

#Robinson's formula
#empirical bayes
empirical <- NULL
for (i in 1:length(counts)){
  empirical[i] <- round(Claims[i+1]*(counts[i+1]/counts[i]),2)
}

empirical<- empirical[1:7]
#gamma-prior

f <- function(x,nu,sigma){
  gamma = sigma / (1 + sigma)
  numer = gamma ^ (nu + x) * gamma(nu + x)
  denom = sigma ^ nu * gamma(nu) * factorial(x)
  return(numer/denom)
}

negloglikelihood <- function(params){
  nu <- params[1]
  sigma <- params[2]
  eqt <- -sum(counts*log(f(Claims, nu=nu, sigma=sigma)))
  return(eqt)
}

p<- as.matrix(c(0.5, 1), ncol=2)
results <- nlm(f = negloglikelihood, p= p, hessian=T)
nu <- results$estimate[1]
sigma<- results$estimate[2]

gamma_mle <- round((seq(0,6)+1)*f(seq(0,6)+1,nu,sigma)/f(seq(0,6),nu,sigma), 2)

table1<- cbind(Claims, counts, empirical, gamma_mle) %>% as.data.frame()

table1$gamma_mle=round(c(f(seq(0,6),nu,sigma)*9461,NA),2)

kable(rbind(Claims, counts, empirical, gamma_mle), "pipe", caption = "European Automobile Insurance")

# log(counts) vs claims for 9461 auto insurance policies. The dashed line is a gamma MLE fit.
ggplot(data=table1)+
  geom_point(aes(x=Claims, y=log(counts)), col="blue")+
  geom_line(aes(x=Claims, y= log(gamma_mle)), col="tomato", linetype = "dashed")+ theme_classic()+ ggtitle("Auto Accident Data")

```


# The Missing-Species Problem

```{r warning=FALSE}
x<- seq(1,24)
y <- c(118, 74, 44, 24, 29, 22, 20, 19, 20, 15, 12, 14, 6, 12, 6, 9, 9, 6, 10, 10, 11, 5, 3, 3)

butterfly <- data.frame(x, y)
t= seq(0, 1, 0.1)

exp <- NULL
sd <- NULL
for (i in 1:length(t)){
  exp[i] <- round(sum(y*(t[i]^x)*(-1)^(x-1)),2)
  sd[i] <- round(sqrt(sum(y*t[i]^(2))),2)
}

table2<- data.frame(t, exp, sd)

```

```{r}
# gamma estimate 
# 0 =< t =<1
v <- 0.104
sigma <-  89.79 
gamma <- sigma / (1 + sigma)
E_1 <- y[1]
gamma_est <- NULL
for (i in 1:length(t)){
  gamma_est[i] <- round(E_1*((1 - (1+gamma*t[i])^(-v)) / (gamma * v)),2)
}

E_1 <- y[1]
gamma_est <- NULL
for (i in 1:length(t)){
  gamma_est[i] <- round(E_1*((1 - (1+gamma*t[i])^(-v)) / (gamma * v)),2)
}

kable(rbind(t, exp, sd, gamma_est), "pipe", caption ="The Missing-Species" )
```

```{r}
# Nonparametric fit (solid) +/- 1 standard deviation; gamma model (dashed).
ggplot(data=table2, aes(x=t))+
  geom_line(aes(y=exp))+
  geom_line(aes(y=gamma_est), col="red", linetype="dashed")+
  geom_errorbar(aes(ymin=(exp-sd), ymax=(exp+sd)), width=0, alpha=0.5)+
  ggtitle("Butterfly Data")+ylab("E(t)")+xlab("Time t") + theme_classic()+
  theme(legend.position="topleft")

```

## Shakespeare’s word counts

```{r warning=FALSE}
data(bardWordCount)
# str(bardWordCount)
lambda <- seq(-4, 4.5, .025)
tau <- exp(lambda)
result <- deconv(tau = tau, y = bardWordCount, n = 100, c0=2)
stats <- result$stats

# Empirical Bayes deconvoluation estimates
ggplot() +
    geom_line(mapping = aes(x = lambda, y = stats[, "g"])) +
    labs(x = expression(log(theta)), y = expression(g(theta)))+theme_classic()

d <- data.frame(lambda = lambda, g = stats[, "g"], tg = stats[, "tg"], SE.g = stats[, "SE.g"]) #tg=correction for truncation at zero
indices <- seq(1, length(lambda), 5)
ggplot(data = d) +
    geom_line(mapping = aes(x = lambda, y = g)) +
    geom_errorbar(data = d[indices, ],
                  mapping = aes(x = lambda, ymin = g - SE.g, ymax = g + SE.g),
                  width = .01, color = "blue") +
    labs(x = expression(log(theta)), y = expression(g(theta))) +
    ylim(0, 0.006) +
    geom_line(mapping = aes(x = lambda, y = tg), linetype = "dashed", color = "red")+theme_classic()

# Posterior estimates
gPost <- sapply(seq_len(100), function(i) local({tg <- d$tg * result$P[i, ]; tg / sum(tg)}))
plots <- lapply(c(1, 2, 4, 8), function(i) {
    ggplot() +
        geom_line(mapping = aes(x = tau, y = gPost[, i])) +
        labs(x = expression(theta), y = expression(g(theta)),
             title = sprintf("x = %d", i))+theme_classic()
})
plots <- Map(f = function(p, xlim) p + xlim(0, xlim), plots, list(6, 8, 14, 20))

plot_grid(plotlist = plots, ncol = 2)
```



## A Medical Example
```{r}
data(surg)
tau <- seq(from = 0.01, to = 0.99, by = 0.01)
result <- deconv(tau = tau, X = surg, family = "Binomial", c0 = 1)
d <- data.frame(result$stats)
indices <- seq(5, 99, 5)
errorX <- tau[indices]

# estimated prior density of $g(\theta)$
ggplot() +
    geom_line(data = d, mapping = aes(x = tau, y = g), alpha=0.5) +
    geom_errorbar(data = d[indices, ], mapping = aes(x = theta, ymin = g - SE.g, ymax = g + SE.g),width =0, color = "red") +
    labs(x = expression(theta), y = expression(paste(g(theta), " +/- SE")))+theme_classic()

# kable(d[indices, ], row.names = FALSE)
```

```{r}
# Posterior Estimates
theta <- result$stats[, 'theta']
gTheta <- result$stats[, 'g']
f_alpha <- function(n_k, x_k) {
    ## .01 is the delta_theta in the Riemann sum
    sum(dbinom(x = x_k, size = n_k, prob = theta) * gTheta) * .01
}
g_theta_hat <- function(n_k, x_k) {
    gTheta * dbinom(x = x_k, size = n_k, prob = theta) / f_alpha(n_k, x_k)
}

# Empirical Bayes posterior densities of $\theta$ for three patients,
# given x= number of positive nodes,n= number of nodes.
g1 <- g_theta_hat(x_k = 7, n_k = 32)
g2 <- g_theta_hat(x_k = 3, n_k = 6)
g3 <- g_theta_hat(x_k = 17, n_k = 18)
ggplot() +
    geom_line(mapping = aes(x = theta, y = g1), col = "skyblue", linetype="dashed") +
    ylim(0, 10) +
    geom_line(mapping = aes(x = theta, y = g2), col = "red") +
    geom_line(mapping = aes(x = theta, y = g3), col = "black", linetype="dotted") +
    labs(x = expression(theta), y = expression(g(paste(theta, "|(x, n)")))) +
    annotate("text", x = 0.15, y = 4.25, label = "x=7, n=32") +
    annotate("text", x = 0.425, y = 4.25, label = "x=3, n=6") +
    annotate("text", x = 0.85, y = 7.5, label = "x=17, n=18")+
  theme_classic()

```

## Main takeaway from this project 

In this project, I learned how to apply empirical Bayes methods to real-world data. However, I struggled a lot because I don't have strong statistical background knowledge. For this semester, I feel like we were doing something different from last semester in which we intended to combine the theoretical part with the empirical part. Over the summer, I plan to continue reading those textbooks and hope I can gain more statistical knowledge and improve my coding skill.

## Acknowledgement 

I would like to thank my classmates, Zening Ye and Shuting Li, who explained the project to me. A special thanks to Yuli Jin who helped me with coding.

## References

[1] Haviland's lecture notes

[2] https://github.com/jrfiedler/CASI_Python/tree/master/chapter06

[3] https://github.com/bnaras/deconvolveR/blob/master/vignettes/deconvolution.Rmd



